Induced universal graphs for families of small graphs

### 2021-10-12 paper/paper.tex
Tweak writing, and add example figure

\usepackage[letterpaper, total={6in, 8in}]{geometry}
\usepackage[x11names, svgnames, rgb]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{amsthm}
\usetikzlibrary{snakes,arrows,shapes}
\usepackage{amsmath}

@@

@@

We write $f(k)$ to denote the order (that is, the
number of vertices) of a minimal induced universal graph for the family of all
graphs on $k$ vertices.  Moon showed that $f(k) \leq 2^{(k-1)/2}$
\cite{moon_1965}, and Alon showed that $f(k) = (1 + o(1))2^{(k-1)/2}$
\cite{alon2017asymptotically}.  There is an extensive literature on bounds on
the order of minimal induced universal subgraphs for many families of graphs;
see the references in \cite{alon2017asymptotically}.
However, to my knowledge the only
existing systematic attempt to find exact values for families of small
graphs is a Mathematics Stack Exchange answer by James Preen that presents results for families of small connected graphs obtained by brute-force search with the Maple programming language \cite{preen_math_se}.

@@

\begin{table}[h!]
\centering
\begin{tabular}{r r r}
 \toprule
 $k$ & $t(k)$ & $T(k)$ \\ [0.5ex]
 \midrule
 0 & 0 & 1 \\
 1 & 1 & 1 \\
 2 & 2 & 1 \\
 3 & 3 & 1 \\
 4 & 5 & 2 \\
 5 & 7 & 18 \\
 6 & 9 & 66 \\
 \bottomrule
\end{tabular}
\caption{For each $k$, $t(k)$ is the minimum order of a graph containing all $k$-vertex trees as
induced subgraphs, and $T(k)$ is the number of distinct $t(k)$-vertex graphs that contain
all $k$-vertex trees as induced subgraphs.}
\label{tab:treeresults}
\end{table}

>>>>

\usepackage[letterpaper, total={6in, 8in}]{geometry}
\usepackage[x11names, svgnames, rgb]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{amsthm}
\usetikzlibrary{decorations, decorations.pathreplacing, decorations.pathmorphing,
    calc, backgrounds, positioning, tikzmark, patterns, matrix, fit, graphs,
    snakes, arrows, shapes, calligraphy}
%\usetikzlibrary{snakes,arrows,shapes}
\usepackage{amsmath}
\usepackage{subfigure}
\usepackage{cleveref}

@@

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}

@@

We write $\calF(k)$ to denote the family of all graphs on $k$ vertices,
and we write $f(k)$ to denote the order (that is, the
number of vertices) of a minimal induced universal graph for $\calF(k)$.
We write $F(k)$ to denote the number of non-isomorphic graphs of order $f(k)$
that are induced universal for $f(k)$.
To give an example, we have $f(3)=5$ and $F(3)=5$. All five of the minimal induced universal graphs
for $\calF(3)$ are shown in \Cref{fig:univ3a}. \Cref{fig:univ3b} shows the four
graphs in $\calF(3)$ as induced subgraphs of a single induced universal graph.

Moon showed that $f(k) \leq 2^{(k-1)/2}$
\cite{moon_1965}, and Alon showed that $f(k) = (1 + o(1))2^{(k-1)/2}$
\cite{alon2017asymptotically}.  There is an extensive literature on bounds on
the order of minimal induced universal subgraphs for many families of graphs;
see the references in \cite{alon2017asymptotically}.
However, to my knowledge the only
existing systematic attempt to find \emph{exact} values for families of small
graphs is an answer by James Preen on the Mathematics Stack Exchange website that
presents results for families of small connected graphs obtained by brute-force
search with the Maple programming language \cite{preen_math_se}.

    \subfigure[][The five induced universal graphs of order 5 for \calF(3) (that is, for the family of all 3-vertex graphs)] {
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1,2,3,4;
           3 -- 0;
           4 -- 0;
           4 -- 1;
           4 -- 3;
        };
      \end{tikzpicture}
      \qquad
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1,2,3,4;
          3 -- 0;
          4 -- 0;
          4 -- 1;
          4 -- 2;
          4 -- 3;
        };
      \end{tikzpicture}
      \qquad
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1,2,3,4;
          3 -- 0;
          3 -- 1;
          4 -- 0;
          4 -- 2;
          4 -- 3;
        };
      \end{tikzpicture}
      \qquad
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1,2,3,4;
          3 -- 0;
          3 -- 1;
          4 -- 0;
          4 -- 1;
          4 -- 3;
        };
      \end{tikzpicture}
      \qquad
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1,2,3,4;
          3 -- 0;
          3 -- 1;
          4 -- 0;
          4 -- 1;
          4 -- 2;
          4 -- 3;
        };
      \end{tikzpicture}
      \label{fig:univ3a}
    }

    \par\bigskip
    \subfigure[][A demonstration that the first graph in \Cref{fig:univ3a} is induced universal
            for $\calF(3)$.  For each graph $G$ in $\calF(3)$ ($I_3$, $K_3$, $P_3$, and
            a graph with a single edge), an induced subgraph isomorphic to $G$ is shown in a
            single color.] {
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1[fill=uofgpumpkin],2[fill=uofgpumpkin],3[fill=uofgpumpkin],4;
           3 -- 0;
           4 -- 0;
           4 -- 1;
           4 -- 3;
        };
      \end{tikzpicture}
      \qquad
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0[fill=uofgturquoise],1,2,3[fill=uofgturquoise],4[fill=uofgturquoise];
           4 -- 1;
           3 --[uofgturquoise, ultra thick] 0;
           4 --[uofgturquoise, ultra thick] 0;
           4 --[uofgturquoise, ultra thick] 3;
        };
      \end{tikzpicture}
      \qquad
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1[fill=uofgrose],2,3[fill=uofgrose],4[fill=uofgrose];
           3 -- 0;
           4 -- 0;
           4 --[uofgrose, ultra thick] 1;
           4 --[uofgrose, ultra thick] 3;
        };
      \end{tikzpicture}
      \qquad
      \begin{tikzpicture}[>=latex',line join=bevel,scale=.4]
        \graph [nodes={draw, circle, minimum width=.52cm, inner sep=1pt}, circular placement, radius=0.82cm,
                clockwise=5] {
                    0,1[fill=uofgmocha],2[fill=uofgmocha],3,4[fill=uofgmocha];
           3 -- 0;
           4 -- 0;
           4 --[uofgmocha, ultra thick] 1;
           4 -- 3;
        };
      \end{tikzpicture}
      \label{fig:univ3b}
    }
\caption{Induced universal graphs for the family of all graphs on 3 vertices}
\label{fig:graphs}
\end{figure}

@@

\begin{table}[h!]
\centering
\begin{tabular}{r r r}
 \toprule
 $k$ & $t(k)$ & $T(k)$ \\ [0.5ex]
 \midrule
 1 & 1 & 1 \\
 2 & 2 & 1 \\
 3 & 3 & 1 \\
 4 & 5 & 2 \\
 5 & 7 & 18 \\
 6 & 9 & 66 \\
 \bottomrule
\end{tabular}
\caption{For each $k$, $t(k)$ is the minimum order of a graph containing all $k$-vertex trees as
induced subgraphs, and $T(k)$ is the number of distinct $t(k)$-vertex graphs that contain
all $k$-vertex trees as induced subgraphs.}
\label{tab:treeresults}
\end{table}
